\fontfamily{phv}\selectfont{Causal Discovery methods for Effective Connectivity in Human Brains}

### 2014-01-14 docs/paper/paper.tex

The results from the agorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ or $n,m$ in the matrix.
For structural data, this matrix is therefore symmetrical.
In the directional data, point $(n,m)$ denotes a directed edge from $n$ to $m$.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum. %TODO left or right hemisphere?
Modified PC and EMS-PC do not differ in finding the structure.
The results are presented in figure ~\ref{fig:struct_avg} as the average over all six subjects, with 300 iterations per subject. Additionally, separate parts of the two hemispheres, the inter-hemisphere, and the cerebellum are presented in Figures \ref{fig:struct_left}, \ref{fig:struct_right}, \ref{fig:struct_inter_hemisphere}, \ref{fig:struct_cerebellum}, \ref{fig:struct_apart}
The results of Hinne et al. are shown as well for comparison.

% TODO: Add in caption what the reader is supposed to be looking at.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/struct_full}
  \includegraphics[width=0.5\textwidth]{images/structure_max}
  \caption{The avarage structure of all subjects. The structure on the left is found with the PC-algorithm, and the structure on the right is found by Hinne et al.}
\label{fig:struct_avg}
\end{figure}
In this matrix, higher values represent strong certainties of the presence of a connection.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.25\textwidth]{images/struct_hemisphere_1_45}
  \includegraphics[width=0.25\textwidth]{images/struct_hemisphere_46_90}
  \includegraphics[width=0.25\textwidth]{images/struct_inter_hemisphere}
  \includegraphics[width=0.25\textwidth]{images/struct_cerebellum}  
  \caption{The average structure of the left and right hemisphere, the inter-hemisphere, and the cerebellum of all subjects, as found with PC.}
  \label{fig:struct_apart}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics{images/arrowheads_avg_mod}
  \caption{The asymmetric arrowheads of the PDAG found with modified PC, in which a high value denotes a high certainty of an one-directional arrow. The average of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_mod}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{images/arrowheads_avg_expl}
  \caption{The asymmetric arrowheads of the PDAG found with EMS-PC, in which a high value denotes a high certainty of an one-directional arrow. The average of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_ems}
\end{figure}

@@

\subsection{Execution time of the algorithm}
To apply the algorithms on our data, we have used a Xeon E5-2670, running on 2,6GHz.
To run one iteration of the PC algorithm for one subject, modified PC needed approximately one minute and EMS-PC needed approximately two minutes to find both structure and direction.
Although PC is not designed to be multithreaded, running multiple iterations lends itself for using multiple threads, as the iterations are fully independent, and only need to be averaged out afterwards.

>>>>

The results from the agorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ or $n,m$ in the matrix.
For structural data, this matrix is therefore symmetrical.
In the directional data, point $(n,m)$ denotes a directed edge from $n$ to $m$.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum. %TODO left or right hemisphere?
Modified PC and EMS-PC do not differ in finding the structure.

%  uit een merge-conflict die ontstaan is bij de haastige inleverpoging van de vorige deadline
%The results are presented in figure ~\ref{fig:struct_avg} as the avarage over all six subjects, with 300 iterations per subject.
%The results of Hinne et al. are shown in ~\ref{fig:struct_max} for comparison.

%\begin{figure}[h!]
%  \centering
%  \includegraphics{images/struct_full_mean_gray}
%  \caption{The avarage structure of all subjects, found with the PC-algorithm}
%  \label{fig:struct_avg}

The results are presented in figure ~\ref{fig:struct_avg} as the average over all six subjects, with 300 iterations per subject. Additionally, separate parts of the two hemispheres, the inter-hemisphere, and the cerebellum are presented in Figures \ref{fig:struct_left}, \ref{fig:struct_right}, \ref{fig:struct_inter_hemisphere}, \ref{fig:struct_cerebellum}, \ref{fig:struct_apart}
The results of Hinne et al. are shown as well for comparison.

% TODO: Add in caption what the reader is supposed to be looking at.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/struct_full}
  \includegraphics[width=0.5\textwidth]{images/structure_max}
  \caption{The avarage structure of all subjects. The structure on the left is found with the PC-algorithm, and the structure on the right is found with DWI-measurements by Hinne et al.}
\label{fig:struct_avg}
\end{figure}
In this matrix, higher values represent strong certainties of the presence of a connection.

@@

%<<<<<<< HEAD
%\begin{figure}[h!]
%  \centering
%  \includegraphics{images/structure_max}
%  \caption{The avarage structure of all subjects, according to the DWI-measurements of Hinne et al.}
%  \label{fig:struct_max}
%\end{figure}

One notable difference between the results found in Hinne et al. is the strong connections between homological areas, visible as two diagonal lines.
This is a known shortcoming of diffusion weighted imaging, as these connections are typically too long to measure correctly.
These connections most likely correspond to the corpus callosum, as each region is connected with its counterpart in the other hemisphere.
Connections within the hemispheres are less clearly visible, as well as connections within the cerebellum.

The second step of the PC-algorithm finds directions based on this structure.
In Figures ~\ref{fig:pdag_avg_mod} and ~\ref{fig:pdag_avg_ems}, the results for both modified PC and EMS-PC are presented.
The figures look very similar to the Figure ~\ref{fig:struct_full_mean}, and look strongly symmetric.
In terms of causality, this means that many latent sources are still present and causal relations between the measured brain regions themselves are not clearly visible.

%\begin{figure}[h!]
%  \centering
%  \includegraphics{images/PDAG_avg_mod}
%  \caption{The PDAG found with modified PC, in which a high value denotes a high certainty of an arrow. The avarage of all six subjects is taken.}
%  \label{fig:pdag_avg_mod}
%=======
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.25\textwidth]{images/struct_hemisphere_1_45}
  \includegraphics[width=0.25\textwidth]{images/struct_hemisphere_46_90}
  \includegraphics[width=0.25\textwidth]{images/struct_inter_hemisphere}
  \includegraphics[width=0.25\textwidth]{images/struct_cerebellum}  
  \caption{The average structure of the left and right hemisphere, the inter-hemisphere, and the cerebellum of all subjects, as found with PC.}
  \label{fig:struct_apart}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics{images/arrowheads_avg_mod}
  \caption{The asymmetric arrowheads of the PDAG found with modified PC, in which a high value denotes a high certainty of an one-directional arrow. The average of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_mod}
%>>>>>>> 6d0d26f13f6a0b4326b7a5c962d4444f78f8d2ea
\end{figure}

\begin{figure}[h!]
  \centering
%<<<<<<< HEAD
%  \includegraphics{images/PDAG_avg_expl}
%  \caption{The PDAG found with EMS-PC, in which a high value denotes a high certainty of an arrow. The avarage of all six subjects is taken.}
%  \label{fig:pdag_avg_ems}
%\end{figure}

%Another representation of this data has been used to emphasize the assymetric parts of Figures ~\ref{fig:pdag_avg_mod} and ~\ref{fig:pdag_avg_ems}.
%To achieve this, the absolute difference between these matrices and their transpose has been calculated, lifting out those points for which there is an arrow in one direction, but not in the opposite direction.
%These results are presented in Figures ~\ref{fig:pdag_avg_antisymmetric_mod} and ~\ref{fig:pdag_avg_antisymmetric_ems}.

%\begin{figure}[h!]
%  \centering
%  \includegraphics{images/PDAG_avg_antisymmetric_mod}
%  \caption{The assymetric parts of the PDAG found with modified PC, in which a high value denotes a high certainty of an one-directional arrow. The avarage of all six subjects is taken.}
%  \label{fig:pdag_avg_antisymmetric_mod}
%\end{figure}
%=======
  \includegraphics{images/arrowheads_avg_expl}
  \caption{The asymmetric arrowheads of the PDAG found with EMS-PC, in which a high value denotes a high certainty of an one-directional arrow. The average of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_ems}
\end{figure}

@@

\subsection{Execution time of the algorithm}
To apply the algorithms on our data, we have used a Xeon E5-2670, running on 2,6GHz.
To run one iteration of the PC algorithm for one subject, modified PC needed approximately one minute and EMS-PC needed approximately two minutes to find both structure and direction.
Although PC is not designed to be multithreaded, running multiple iterations lends itself for using multiple threads, as the iterations are fully independent, and only need to be averaged out afterwards.
%>>>>>>> 6d0d26f13f6a0b4326b7a5c962d4444f78f8d2ea

\begin{figure}[h!]
  \centering
  \includegraphics{images/PDAG_avg_antisymmetric_expl}
  \caption{The assymetric parts of the PDAG found with EMS-PC, in which a high value denotes a high certainty of an one-directional arrow. The avarage of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_ems}
\end{figure}
